Model Based Analysis of Patient Derived Xenograft Growth Curves

### 2022-05-01 paper_Kuhlers.Rmd
figure edits

Tumor volume data was obtained from the PDXnet Portal,
queried for all available colon adenocarcinoma.
All data were contributed from the NCI Patient-Derived Models Repository (PDMR)[@noauthor_patient-derived_nodate].
Mice were staged to a tumor volume of ~200mm^3^ before being randomized to a treatment group.
Tumors were measured by caliper 1-3 times per week,
potentially more depending on the health of animal[@evrard_systematic_2020; @noauthor_patient-derived_nodate; @teicher_human_2004].
The exact study endpoints for this data were not documented,
however Evrard et al performed a similar study with data from the PDMR,
where the study endpoints were "Animal health, Max. tumor size, 4000mm^3^,
or 300 days, if max. TV not reached" (See Table 1 of Evrard et al[@evrard_systematic_2020]).
Additionally, the PDMR standard operating procedures
recommend harvesting tumors at 1000-2000mg[@noauthor_patient-derived_nodate].
The distribution of maximum tumor volumes in this dataset reflects this.
Observation days were rescaled to the number of days the sample
was observed, so that the time origin was 0 for all samples.
For example, if the mouse was staged on observation day 30,
this became day 0, and all subsequent measurements on this mouse were scaled to
days since day 30. Days were then rescaled to months.
All data processing and analysis were conducted using R/v4.1.0 ([@r-core]).

@@

\begin{figure}
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 5in]{figures/growth_rescaled_days.pdf}}
\caption{Log tumor volumes over time.
Black lines indicate each mouses' growth curve.
Colored lines are simple linear regressions of log-volume versus time,
separated by each treatment group}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 5in]{figures/mixed_forest.pdf}}
\caption{Forest plot of exponentiated mixed effects model interaction coefficients.
The coefficients represent the change in the ratio of mean volume in treatment versus control
over time. Red points indicate the growth rate is reduced in the treatment group,
while blue points indicate the growth rate is increased in the treatment group.
Horizontal bars indicate 95\% confidence intervals.}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 5in]{figures/aft_models_overlaid_forest.pdf}}
\caption{Forest plot of AFT acceleration factors.
These represent the ratio of the mean time to 1500mm$^3$ in treatment versus control.
Red points indicate that time to 1500mm$^3$ is extended relative to control,
while blue points indicate the time is shortened.}
\end{center}
\end{figure}

>>>>

Tumor volume data was obtained from the PDXnet Portal,
queried for all available colon adenocarcinoma.
All data were contributed from the NCI Patient-Derived Models Repository (PDMR)[@noauthor_patient-derived_nodate].
Mice were staged to a tumor volume of ~200mm^3^ before being randomized to a treatment group.
Tumors were measured by caliper 1-3 times per week,
potentially more depending on the health of animal[@evrard_systematic_2020; @noauthor_patient-derived_nodate; @teicher_human_2004].
The exact study endpoints for this data were not documented,
however Evrard et al performed a similar study with data from the PDMR,
where the study endpoints were "Animal health, Max. tumor size, 4000mm^3^,
or 300 days, if max. TV not reached" (See Table 1 of Evrard et al[@evrard_systematic_2020]).
Additionally, the PDMR standard operating procedures
recommend harvesting tumors at 1000-2000mg[@noauthor_patient-derived_nodate].
The distribution of maximum tumor volumes in this dataset reflects this.
Observation days were rescaled to the number of days the sample
was observed, so that the time origin was 0 for all samples.
For example, if the mouse was staged on observation day 30,
this became day 0, and all subsequent measurements on this mouse were scaled to
days since day 30. Days were then rescaled to months.
All data processing and analysis were conducted using R/v4.1.0[@r-core].

@@

\begin{figure}[H]
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 4.5in]{figures/growth_rescaled_days.pdf}}
\caption{Log tumor volumes over time.
Black lines indicate each mouses' growth curve.
Colored lines are simple linear regressions of log-volume versus time,
separated by each treatment group}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 4.5in]{figures/mixed_forest.pdf}}
\caption{Forest plot of exponentiated mixed effects model interaction coefficients.
The coefficients represent the change in the ratio of mean volume in treatment versus control
over time. Red points indicate the growth rate is reduced in the treatment group,
while blue points indicate the growth rate is increased in the treatment group.
Horizontal bars indicate 95\% confidence intervals.}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\shadowbox{\includegraphics[width = 6in, height = 4.5in]{figures/aft_models_overlaid_forest.pdf}}
\caption{Forest plot of AFT acceleration factors.
These represent the ratio of the mean time to 1500mm$^3$ in treatment versus control.
Red points indicate that time to 1500mm$^3$ is extended relative to control,
while blue points indicate the time is shortened.}
\end{center}
\end{figure}
